---
title: "DPS Covid"
subtitle: "Can you wrangle the dashboard?"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )

# load packages
  library(tidyverse)
```


Durham Public Schools publishes a COVID-19 "dashboard". Check it out [here](https://www.dpsnc.net/Page/5339). 

Click on any of the links and you will find a PDF. There are tools for extracting data from pdfs, but fortunately you'll find a [link to a spreadsheet](https://docs.google.com/spreadsheets/d/1kQ35JKpk3vNaM7ZfFK2DFb1B9kQIQ4IODr7lOM394nQ/edit#gid=769268401). Please open it.

## 1. Describe the data structure. 

The spreadsheet records the number of positive COVID-19 cases by week, as well as the total number of cumulative cases since August 2021. The data are disaggregated by: week, school, central services/elementary/middle/high, and staff/students/clusters. Each school appears to have its own unique ID in the column "Code."

## 2. This sheet appears to be updated weekly. What are some barriers to feeding this sheet into a data pipeline that will always plot the latest results?

Date is not in its own column, so case data are not associated with dates except visually in the spreadsheet. That is, cases for a given week are in three columns headed by the name of that week.

## 3. Import this sheet into R and assign it to an object called `raw`. (hint: you need to do this in a way that your analysis will update every time you knit this document)

```{r}
library(googlesheets4)
gs4_deauth()
raw <- read_sheet("https://docs.google.com/spreadsheets/d/1kQ35JKpk3vNaM7ZfFK2DFb1B9kQIQ4IODr7lOM394nQ/edit#gid=769268401") 
```

## 4. What are the steps you need to do to get this into a tidy dataframe?

I want to get each week's observations into its own row. For each week, I want different columns for students, staff, clusters. The unique identifier for each row will be the combination of id-week.

code | name      | week       | cases_students | cases_staff | cases_clusters
xx   | xx        | YYYY-MM-DD | xx             | xx          | xx

Let's tackle student, staff and clusters one at a time. Student cases are currently wide and I want them to go long. This requires each of my student columns to be labeled with the date. Assuming I do that, the next step is:
pivot_longer(c(`2022-01-23`, `2022-01-16`), names_to = "week", values_to = "cases_students")

## 5. Go!

```{r}
tidyraw <- raw %>%
  select(-c(3:5)) # get rid of the cumulative columns
  
           
date_vector <- tidyraw %>%
  slice(2) %>%
  unlist(use.names = FALSE) %>%
  as_tibble() %>%
  mutate(value = str_remove(value, "Positive COVID-19 Cases\n")) %>%
  mutate(value = str_trim(value, "both"))

# first, fix strings
  
# str_sub()
# rep(date_vector[c(3, 28)], each=3)

```

