---
title: "Data Wrangling"
subtitle: "Data transformation"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )

# load packages
  library(nycflights13)
  library(tidyverse)
```

# `filter()`

`filter()` allows you to subset observations based on their values 

## Find all flights that had an arrival delay of two or more hours

```{r}
  flights %>% 
    filter()
```

## Find all flights that flew to Houston (IAH or HOU)

```{r}

```

## Find all flights that were operated by United, American, or Delta, but don't use the OR operator (`|`)

```{r}

```

## Find all flights that departed in summer (July, August, and September) using `between()`

```{r}

```

# `arrange()`

`arrange()` sorts the observations

## How could you use arrange(dep_time) to sort all missing values to the start?

`arrange()` always puts missing values at the end, even if you use `desc()`. The trick is to use `is.na()` because this evaluates to TRUE or FALSE. TRUE > FALSE in the default sorting.

```{r}
# these put NA last
  flights %>% arrange(dep_time)
  flights %>% arrange(desc(dep_time))
  
# make NA come first
  
```

## Sort flights to find the flights with the shortest flight time

```{r}

```

# `select()`

`select()` subsets the data by column

## What does the `any_of()` function do? Why might it be helpful in conjunction with this vector? How is it different from `all_of()`?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

flights %>% 
  select()
```

# `mutate()`

`mutate()` adds new columns (or overwrites old ones) that are functions of existing columns

## Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

Jeffrey Arnold has a nice suggestion for working through this one:

> The dplyr package provides multiple functions for ranking, which differ in how they handle tied values: row_number(), min_rank(), dense_rank(). To see how they work, let’s create a data frame with duplicate values in a vector and see how ranking functions handle ties.

```{r}
tibble(x = c(10, 5, 1, 5, 5)) %>%
  mutate(
    x_row_number = row_number(x),
    x_min_rank = min_rank(x),
    x_dense_rank = dense_rank(x)
  ) %>%
  arrange(x)
```

How could you use `arrange()` and `slice()`? Or `top_n()`? What about ties?

```{r}
flights %>%
  arrange() %>%
  slice()

flights %>%
  arrange() %>%
  slice()
```
# `summarize()`

`summarize()` collapses a data frame to a single summary row

## Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)

```{r}
flights %>% 
  group_by(carrier, dest) %>% 
  summarise(n())
```

Carriers can experience days related to the airports they fly into and out of. One approach is to look at delays by route.

```{r}
flights %>%
# calculate delays by route and carrier
  filter() %>%    # ?? look just at records with data for arr_delay
  group_by() %>%  # ?? we want to look at combinations of carrier and route
  summarise(
    arr_delay = sum(arr_delay),
    flights = n()
  ) %>%
# calculate total delay by route
  group_by() %>%  # ??
  mutate(
    arr_delay_total = sum(arr_delay),
    flights_total = sum(flights)
  ) 
```
If you fill in the blanks correctly you'll be part of the way to an answer. What are the remaining things you need to do?



