---
title: "Communicating Data Science"
subtitle: "RMarkdown"
css: custom_report.css
bibliography: refs.bib
csl: nature.csl
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r example, include=FALSE, warning = FALSE, message = FALSE}
```

```{r setup}
#| include = FALSE,
#| warning = FALSE,
#| message = FALSE
  
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )

# load packages
  library(tidyverse)
  library(gapminder)
```

# Reproducible Research

* Goals per @peng:2021
  * The original goal was to better reveal the process of doing the research.
  * Some have suggested that making data and computer code available to others is a per se goal of reproducible research because both can be built on and leveraged to further scientific knowledge.
  * Another goal of reproducible research is to provide a kind of audit trail, should one be needed.
  * One could summarize the goal of reproducible research as providing a means to answer the question, "Do I understand and trust this data analysis?"



# References




