---
title: "week03_class1"
author: "tamsin-chen"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

# Week 3, class 1

## R4DS Chapter 8

Save R script as diamonds.R

```{r save PDFs and CSVs straight from R}
ggplot(diamonds, aes(carat, price)) + 
  geom_hex()
ggsave("diamonds.pdf")

write_csv(diamonds, "diamonds.csv")
```

## R4DS Chapter 10

Tibbles are part of the core of the tidyverse package.
Tibbles are data frames, but they tweak some of the older behavior of normal data frames in R.

```{r coerce data into tibbles}
as_tibble(iris)
#> # A tibble: 150 x 5
#>   Sepal.Length Sepal.Width Petal.Length Petal.Width Species
#>          <dbl>       <dbl>        <dbl>       <dbl> <fct>  
#> 1          5.1         3.5          1.4         0.2 setosa 
#> 2          4.9         3            1.4         0.2 setosa 
#> 3          4.7         3.2          1.3         0.2 setosa 
#> 4          4.6         3.1          1.5         0.2 setosa 
#> 5          5           3.6          1.4         0.2 setosa 
#> 6          5.4         3.9          1.7         0.4 setosa 
#> # … with 144 more rows
```

```{r create tibble manually from vectors}
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
#> # A tibble: 5 x 3
#>       x     y     z
#>   <int> <dbl> <dbl>
#> 1     1     1     2
#> 2     2     1     5
#> 3     3     1    10
#> 4     4     1    17
#> 5     5     1    26
```

```{r compare and contrast tibbles and data frames}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

# The two main differences between tibbles and data frames are (1) HOW THEY PRINT and (2) HOW THEY SUBSET DATA.
```

```{r transposed tibbles, or tribbles}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
#> # A tibble: 2 x 3
#>   x         y     z
#>   <chr> <dbl> <dbl>
#> 1 a         2   3.6
#> 2 b         1   8.5
```

## Importing / reading in data

* read_csv() reads comma delimited files
* read_csv2() reads semicolon separated files (common in countries where , is used as the decimal place)
* read_tsv() reads tab delimited files
* read_delim() reads in files with any delimiter
* read_csv(".csv", skip = 2) skips two rows of data (for headers etc. you don't want)
* read_csv(".csv", colnames = FALSE) for data with no column names; R will automatically label them X1-Xn
* read_csv(".csv", na = ".") tells R how N/A values are represented in the csv file you are loading in (in this case, periods)

[Cheat sheet](file:///Users/macbook4/Downloads/data-import.pdf)

