---
title: "week02_class2"
author: "tamsin-chen"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
library(ggokabeito)
library(ggridges)

# If using Google Fonts, also load in library(sysfonts) and then pick your font, e.g., font_add_google("Alegraya Sans", "aleg")
```

# Week 2, class 2

## Walkthrough of Nicci Potts' code
* font_add_google()
* {r fig.width=12.25, fig.height=18}

## SOCVIZ Chapter 4 ggplot

Every geom has an associated stat function that it uses by default. The reverse is also true. Consider geom_bar(), which executes a count function behind the scenes. geom_bar() is associated with stat_count().

geom_bar() can also do relative frequencies, or proportions. That's the ..prop.. stat. But when we use ..prop.., we need to tell ggplot to ignore the x-categories when calculating the denominator of proportions. To do that, we specify group = 1 inside aes(). The value of 1 is just a dummy group that tells ggplot to use the whole dataset when calculating the denominator for proportions.

p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1)) 

## Kieran Healy - Comparing Distributions

* geom_histogram vs. geom_density
* Neat trick using ridgeline plots via geom_density_ridges() from the ggridges package

```{r faceted line and smooth line plots}
p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(color="gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) +
    scale_y_log10(labels=scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year",
         y = "GDP per capita",
         title = "GDP per capita on Five Continents", 
         subtitle = "Individual countries shown in gray, trend in blue.")
```

```{r prep data for subpopulations}
## Keep track of labels for as_labeller() functions in plots later.
grp_names <- c(`a` = "Group A",
               `b` = "Group B",
               `c` = "Group C",
               `pop_a` = "Group A",
               `pop_b` = "Group B",  
               `pop_c` = "Group C",  
               `pop_total` = "Total",                 
               `A` = "Group A", 
               `B` = "Group B", 
               `C` = "Group C")

# make it reproducible
set.seed(1243098)

# 3,000 "counties"
N <- 3e3

# Means and standard deviations of groups
mus <- c(0.2, 1, -0.1)
sds <- c(1.1, 0.9, 1)
grp <- c("pop_a", "pop_b", "pop_c")

# Make the parameters into a list
params <- list(mean = mus, 
                sd = sds)

# Feed the parameters to rnorm() to make three columns, 
# switch to rowwise() to take the average of the columns for
# each row.
df <- pmap_dfc(params, rnorm, n = N) %>% 
  rename_with(~ grp) %>%
  rowid_to_column("unit") %>% 
  rowwise() %>% 
  mutate(pop_total = mean(c(pop_a, pop_b, pop_c))) %>% 
  ungroup()
df

my_oka <- palette_okabe_ito(order = c(1, 2, 3, 5, 7), 
                            alpha = NULL)

df %>%
  pivot_longer(cols = pop_a:pop_total) %>%
  ggplot() + 
  geom_density(mapping = aes(x = value, 
                          color = name, fill = name), 
            alpha = 0.5) + 
  scale_color_manual(values = alpha(c( my_oka[1:3], "gray40"), 1),
                     labels = as_labeller(grp_names)) + 
  scale_fill_manual(values = alpha(c( my_oka[1:3], "gray40"), 0.6),
                    labels = as_labeller(grp_names)) + 
  labs(x = "Measure", y = "Density", 
       title = "Comparing Subgroups: Density", 
       color = "Group", 
       fill = "Group")
```

```{r ridgeline plots}
df %>%
  pivot_longer(cols = pop_a:pop_total) %>%
  ggplot() + 
  geom_density_ridges(mapping = aes(x = value, 
                                    y = name, 
                                    fill = name), 
                      color = "white") + 
  scale_fill_manual(values = alpha(c(my_oka[1:3], "gray40"), 0.7)) + 
  scale_y_discrete(labels = as_labeller(grp_names)) + 
  guides(color = "none", fill = "none") + 
  labs(x = "Measure", y = NULL, title = "Comparing Total and Subgroups: Ridgelines") + 
  theme_ridges(font_family = "")
```

```{r faceted histograms with a reference group}
df %>% 
   pivot_longer(cols = pop_a:pop_c) %>%
   ggplot() + 
   geom_histogram(mapping = aes(x = value, y = ..density.., 
                           color = name, fill = name), 
             stat = "bin", bins = 20, size = 0.5,
             alpha = 0.7) + 
   geom_step(mapping = aes(x = pop_total, y = ..density..), 
                 bins = 20, alpha = 0.9,
                 color = "gray30", size = 0.6, 
             stat = "bin",
             direction = "mid") + 
   scale_fill_manual(values = alpha(my_oka, 0.8)) + 
   scale_color_manual(values = alpha(my_oka, 1)) + 
   guides(color = "none", fill = "none") + 
   labs(x = "Measure", y = "Density", 
        title = "Comparing Subgroups: Histograms", 
        subtitle = "Overall distribution shown in outline") + 
   facet_wrap(~ name, nrow = 1, labeller = as_labeller(grp_names)) 
```

```{r frequency polygons}
#regular
df %>% 
  pivot_longer(cols = pop_a:pop_c) %>%
  ggplot() + 
  geom_area(mapping = aes(x = value, y = ..count.., 
                          color = name, fill = name), 
            stat = "bin", bins = 20, size = 0.5) + 
  geom_freqpoly(mapping = aes(x = pop_total, y = ..count..), 
                bins = 20, 
                color = "gray20", size = 0.5) + 
  scale_fill_manual(values = alpha(my_oka, 0.7)) + 
  scale_color_manual(values = alpha(my_oka, 1)) + 
  guides(color = "none", fill = "none") + 
  labs(x = "Measure", y = "Count", 
       title = "Comparing Subgroups: Frequency Polygons", 
       subtitle = "Overall distribution shown in outline") + 
  facet_wrap(~ name, nrow = 1, labeller = as_labeller(grp_names))

# with scaled counts
df %>% 
  pivot_longer(cols = pop_a:pop_c) %>%
  ggplot() + 
  geom_area(mapping = aes(x = value, y = ..ncount.., 
                          color = name, fill = name), 
            stat = "bin", bins = 20, size = 0.5) + 
  geom_freqpoly(mapping = aes(x = pop_total, y = ..ncount..), 
                bins = 20, 
                color = "gray20", size = 0.5) + 
  scale_fill_manual(values = alpha(my_oka, 0.7)) + 
  scale_color_manual(values = alpha(my_oka, 1)) + 
  guides(color = "none", fill = "none") + 
  labs(x = "Measure", y = "Count", 
       title = "Comparing Subgroups: Frequency Polygons (Scaled Counts)", 
       subtitle = "Overall distribution shown in outline") + 
  facet_wrap(~ name, nrow = 1, labeller = as_labeller(grp_names)) 
```

```{r kernel densities}

#regular
df %>% 
  pivot_longer(cols = pop_a:pop_c) %>%
  ggplot() + 
  geom_density(mapping = aes(x = value, 
                          color = name, fill = name), 
            size = 0.5) + 
  geom_density(mapping = aes(x = pop_total), 
                color = "gray20", size = 0.5) + 
  scale_fill_manual(values = alpha(my_oka, 0.7)) + 
  scale_color_manual(values = alpha(my_oka, 1)) + 
  guides(color = "none", fill = "none") + 
  labs(x = "Measure", y = "Density", 
       title = "Comparing Subgroups: Kernel Densities", 
       subtitle = "Overall distribution shown in outline") + 
  facet_wrap(~ name, nrow = 1, labeller = as_labeller(grp_names)) 


# scaled kernel densities
df %>% 
  pivot_longer(cols = pop_a:pop_c) %>%
  ggplot() + 
  geom_density(mapping = aes(x = value, 
                             y = ..ndensity..,
                          color = name, fill = name), 
            size = 0.5) + 
  geom_density(mapping = aes(x = pop_total, 
                             y = ..ndensity..), 
                color = "gray20", size = 0.5) + 
  scale_fill_manual(values = alpha(my_oka, 0.7)) + 
  scale_color_manual(values = alpha(my_oka, 1)) + 
  guides(color = "none", fill = "none") + 
  labs(x = "Measure", y = "Density", 
       title = "Comparing Subgroups: Kernel Densities (Scaled)", 
       subtitle = "Overall distribution shown in outline") + 
  facet_wrap(~ name, nrow = 1, labeller = as_labeller(grp_names)) 
```

## "Learning to use the patchwork R package" by Riffomonas Project

* CRAN is the official repository of all R packages!
* patchwork composes an array of plots together using the + sign.
* patchwork is a nice alternative to cowplot.
* patchwork defaults to plotting things in a square.
* operators are very important in patchwork. &, *, +, |, \ all mean specific things.
