---
title: "week02_class1"
author: "tamsin-chen"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Week 2, class 1

## Trying out geoms on my own

Search Twitter again for more #TidyTuesday Week 2 (bees) entries. Find examples of at least 3 different geoms used in #TidyTuesday charts and write down the names of these geoms.

From [ocean-v](https://github.com/ocean-v/my_tidytuesday/blob/main/2022_Week_02_code.r) [original twitter post](https://twitter.com/pat_macro/status/1483457396572037124)):
* geom_sf visualizes sf objects. An sf object is a collection of simple features that includes attributes and geometries in the form of a dataframe.
* geom_scatterpie was not in the ggplot2 geom dictionary. Someone developed a scatter pie plot package that plots multiple pie charts in a geometric arrangement.

From [NicciPotts](https://github.com/NicciPotts/tidytuesday/blob/main/bees.R) ([original twitter post](https://twitter.com/nicci_potts/status/1481713459939229704)):
* geom_polygon to create hexagonal honeycomb-like shapes with a fill.
* geom_text is useful for labeling plots (e.g., labeling data points or annotating the height of bars). geom_text adds only text; geom_label adds a rectangle behind the text.
* geom_segment draws a straight line between (x, y) and (xend, yend).

From [efranke22](https://github.com/efranke22/tidytuesday/blob/main/2022/bees_week2.Rmd) ([original twitter post](https://twitter.com/efranke7282/status/1481039935834570755)):
* geom_hline adds reference lines to a plot; useful for annotating.
* geom_curve draws a curved line.
* geom_jitter is a convenient shortcut for geom_point(position = "jitter"). It adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets.

Load in data:
```{r}
tuesdata <- tidytuesdayR::tt_load('2022-01-11')
colony <- tuesdata$colony
stressor <- tuesdata$stressor
```

Prep data:
```{r}
# removing relevant missing values from colony
colony <- colony %>% filter(!is.na(colony_lost))
colony <- colony %>% filter(!is.na(colony_added))
colony <- colony %>% filter(colony_n<10000)
```

Try out geoms:
```{r}
# geom_jitter()
plot1<- ggplot(colony, aes(x = colony_lost, y = colony_added)) +
  geom_jitter()

plot1

# geom_segment()
plot2 <- plot1 +
  geom_segment(aes(x = 5000, y = 5000, xend = 10000, yend = 10000))

plot2

# geom_hline() and geom_vline()
plot3 <- plot2 +
  geom_hline(yintercept = 2000) +
  geom_vline(xintercept = 2000)

plot3
```

## ggplot2 tutorial

Check version of ggplot2 and load it in.
```{r}
packageVersion("ggplot2")
library(ggplot2)
```

Check out the preloaded mtcars dataframe.
```{r}
# see first six lines
head(mtcars)
# see structure of data
str(mtcars)
```

Boxplots
```{r}
ggplot(data = mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_boxplot()
# The x-axis variable is usually a categorical variable for boxplots. We turn cyl into a categorical variable using factor().
```

Mapping aesthetics to continuous variables
```{r}
# Aesthetics are usually mapped to categorical variables (e.g., linetype = stressor, or shape = stressor). But you can map color and size to continuous variables.
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()
```

Mapping aesethetics globally, for specific geoms, and inside/outside of aesthetics
```{r}
# Generally, if you are mapping an aesthetic to a variable from your dataset, you do it inside aes(), but if you are setting an aesthetic to a constant value, you do it outside aes().
ggplot(mtcars, aes(x = factor(cyl), y = mpg, color = factor(cyl))) +
    geom_boxplot() +
    geom_point(color = "purple")
```

Layering geoms with transparency
```{r}
# The last geom coded is the one that lies on top of the whole plot.
# You can change the alpha value to make geoms transparent so they are more visible on top of one another.
ggplot(mtcars, aes(x = mpg, fill = am) ) +
    geom_histogram() + 
    geom_density(aes(y = after_stat(count) ), alpha = 0.2)
```

Facets
Have to fix something here re: manipulating the mtcars dataframe manually first before plotting.
```{r}
# Setting different x-axis scales for different facets.
ggplot(mtcars, aes(x = wt, y = mpg) ) +
    geom_point() +
    facet_wrap(~numcyl, scales = "free_x")
```

## Tidy Tuesday tips
* eval false! Allows you to knit the .rmd by skipping over a piece of code that you know is not going to run.
* Grading is flexible... An 'A' looks different for different people.


