---
title: "DPS Covid"
subtitle: "Can you wrangle the dashboard?"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )

# load packages
  library(tidyverse)
```


Durham Public Schools publishes a COVID-19 "dashboard". Check it out [here](https://www.dpsnc.net/Page/5339). 

Click on any of the links and you will find a PDF. There are tools for extracting data from pdfs, but fortunately you'll find a [link to a spreadsheet](https://docs.google.com/spreadsheets/d/1kQ35JKpk3vNaM7ZfFK2DFb1B9kQIQ4IODr7lOM394nQ/edit#gid=769268401). Please open it.

## 1. Describe the data structure. 



## 2. This sheet appears to be updated weekly. What are some barriers to feeding this sheet into a data pipeline that will always plot the latest results?



## 3. Import this sheet into R and assign it to an object called `raw`. (hint: you need to do this in a way that your analysis will update every time you knit this document)

```{r}

```

## 4. What are the steps you need to do to get this into a tidy dataframe?



## 5. Go!

```{r}

```

