---
title: "Data Visualization"
subtitle: "Common geoms and plot styles to know"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r setup}
#| include = FALSE, 
#| warning = FALSE, 
#| message = FALSE

  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  
# load packages
  library(tidyverse)
  library(palmerpenguins)
  library(viridis)
  library(ggdist)
```

```{r design, include=FALSE}
# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )
  
  options(scipen=999) # turns off scientific notation
```
 

# Histograms and density plots

```{r histograms}
  ggplot(data = penguins, aes(x = flipper_length_mm)) +
    geom_histogram(aes(fill = species),
                   alpha = 0.5,
                   position = "identity") +
    scale_fill_manual(values = c("darkorange","purple","cyan4")) +
    theme_minimal() + # a built-in function that makes your plots look stylistically like this
    labs(x = "Flipper length (mm)",
         y = "Frequency",
         title = "Penguin flipper lengths")

# when you want to fill something up with color, use "fill."
# when you want just the outline, use "color."
```

```{r histograms_yourturn}

# install.packages("viridis")
# load viridis package
library(viridis)

  ggplot(data = penguins, aes(x = flipper_length_mm)) +
    geom_density(aes(fill = species),
                   alpha = 0.8,
                   position = "identity") +
    scale_fill_viridis_d() +
    theme_minimal() +
    labs(x = "Flipper length (mm)",
         y = "Frequency",
         title = "Penguin flipper lengths") +
    theme(legend.position = "bottom")
```


# Violin plots

```{r violin}
  ggplot(data = penguins, aes(x=species, y = flipper_length_mm, fill=species)) +
    geom_violin() +
    scale_fill_manual(values = c("darkorange","purple","cyan4")) +
    theme_minimal() +
    labs(y = "Flipper length (mm)",
         x = NULL,
         title = "Penguin flipper lengths") + 
    theme(legend.position = "none")
```

```{r violin_yourturn}
  ggplot(data = penguins, aes(x=species, y = flipper_length_mm, fill=species)) +
    geom_violin(scale = "count", adjust = 0.5) +
    scale_fill_manual(values = c("darkorange","purple","cyan4")) +
    theme_minimal() +
    labs(y = "Flipper length (mm)",
         x = NULL,
         title = "Penguin flipper lengths") + 
    theme(legend.position = "none") +
  geom_jitter() +
  geom_boxplot(width = 0.1, color = "#FFC300", alpha = 0.8) +
  coord_flip()

# in geom_violin(), if scale = "area," then area of the violin is scaled proportionally to number of observations. if "area," then all violins have the same area.

# notice how in geom_boxplot(), the specifications are not inside aes(). This is because they are constants you are applying.


```

# Raincloud

[Cédric Scherer](https://z3tt.github.io/OutlierConf2021/) shows off a bit with this custom raincloud plot. You don't need to do anything here or get into the weeds.

```{r}
df_peng_stats <- 
  penguins %>% 
  mutate(bill_ratio = bill_length_mm / bill_depth_mm) %>% 
  filter(!is.na(bill_ratio)) %>% 
  group_by(species) %>% 
  mutate(
    n = n(),
    median = median(bill_ratio),
    max = max(bill_ratio)
  ) %>% 
  ungroup() %>% 
  mutate(species_num = as.numeric(fct_rev(species)))

ggplot(df_peng_stats, aes(bill_ratio, species_num, color = species)) +
  stat_summary(
    geom = "linerange",
    fun.min = function(x) -Inf,
    fun.max = function(x) median(x, na.rm = TRUE),
    linetype = "dotted",
    orientation = "y",
    size = .7
  ) +
  geom_point(
    aes(y = species_num - .15), 
    shape = "|",
    size = 5,
    alpha = .33
  ) +
  ggdist::stat_halfeye(
    aes(
      y = species_num,
      color = species,
      fill = after_scale(colorspace::lighten(color, .5))
    ),
    shape = 18,
    point_size = 3,
    interval_size = 1.8,
    adjust = .5,
    .width = c(0, 1)
  ) +
  geom_text(
    aes(x = median, label = format(round(median, 2), nsmall = 2)),
    stat = "unique",
    color = "white",
    family = "Open Sans",
    fontface = "bold",
    size = 3.4,
    nudge_y = .15
  ) +
  geom_text(
    aes(x = max, label = glue::glue("n = {n}")),
    stat = "unique",
    family = "Open Sans",
    fontface = "bold",
    size = 3.5,
    hjust = 0,
    nudge_x = .01,
    nudge_y = .02
  ) +
  coord_cartesian(clip = "off", expand = FALSE) +
  scale_x_continuous(
    limits = c(1.6, 3.8),
    breaks = seq(1.6, 3.8, by = .2)
  ) +
  scale_y_continuous(
    limits = c(.55, NA),
    breaks = 1:3,
    labels = c("Gentoo", "Chinstrap", "Adélie")
  ) +
  scale_color_manual(values = c("#3d6721", "#a86826", "#006c89"), guide = "none") +
  scale_fill_manual(values = c("#3d6721", "#a86826", "#006c89"), guide = "none") +
  labs(
    x = "Bill ratio",
    y = NULL,
    title = "Raincloud plots can tell us so much more",
    subtitle = "Distribution of bill ratios, estimated as bill length divided by bill depth.",
    caption = "Data: Gorman, Williams & Fraser (2014) *PLoS ONE*; Illustration: Allison Horst."
  ) +
  theme_minimal(base_size = 12, base_family = "Open Sans") +
  theme(
    panel.grid.major.x = element_line(size = .35),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 13),
    axis.ticks.length = unit(0, "lines"),
    plot.title.position = 'plot',
    plot.margin = margin(10, 25, 10, 25),
    axis.ticks = element_line(color = "grey92"),
    panel.grid.minor = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(color = "grey30"),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey30"),
    plot.caption = element_text(size = 9, margin = margin(t = 15))
  )
```

# Bar plots

Start by counting penguins for each species / island

```{r}
  penguins %>%
    count(species, island, .drop = FALSE)
```

You have two choices for bar plots: 

1. Let R summarize for you with `geom_bar()`
2. Pre-summarize the data then pass to `geom_col()`

```{r bar}
# geom_bar
  ggplot(penguins, aes(y = species, fill = species)) +
    geom_bar(alpha = 0.8) +
    scale_fill_viridis_d() +
    theme_minimal()
```

geom_bar by default does our count summary for us by species.

You can also pass a weight aesthetic to it. When you plot geom_bar(aes(weight = flipper_length_mm)), you are just summing up flipper lengths by species (this doesn't really make sense as a statistic - just for demonstration).

```{r bar_yourturn}

```

By contrast to geom_bar, geom_col doesn't do anything to your data. It just plots it.

QUICK TIP: Use na.rm = TRUE to remove missing data.

```{r col}
# geom_col
  penguins %>% 
    group_by(species) %>%
    summarize(flip_m = mean(flipper_length_mm, na.rm=TRUE)) %>%
    ggplot(aes(x = species, y = flip_m)) +
      geom_col() + theme_minimal()
```

# Dot plots

```{r dot}
penguins %>%
  remove_missing() %>% group_by(species) %>% # function to remove the missing data
  summarise(mean_bmg = mean(body_mass_g)) %>% # summarize body mass using mean() by species
  ggplot() +
    geom_segment(aes(x = 0, xend = mean_bmg, # I want each line to go from 0 to the mean for that group
                     y = reorder(species, mean_bmg),
                     yend = reorder(species, mean_bmg)), # I want my y values to stay constant
                 color = "grey", size = 2) +
    geom_point(aes(y = reorder(species, mean_bmg), x = mean_bmg),
               size = 5, color = "darkorange") +
    labs(x = NULL, y = NULL, title = "Gentoos are big birds", 
         subtitle = "Average body mass (g) by species") +
    theme_minimal() +
    theme(plot.title = element_text(face="bold"),
          plot.title.position = "plot")
```

```{r dot_yourturn}

```

# Scatterplots and smoothing

```{r scatter}
ggplot(data = penguins,
       aes(x = flipper_length_mm, y = bill_length_mm)) +
  geom_point(aes(color = species),
             size = 3, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper and bill length",
       subtitle = "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)", y = "Bill length (mm)",
       color = "Penguin species") +
  theme(legend.position = "bottom",
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

# There's also a black and white theme: theme_bw(). This makes it look more like a journal plot.
# Investigate why font face for plot.caption isn't working.
```

```{r scatter_yourturn}
ggplot(data = penguins,
       aes(x = flipper_length_mm, y = bill_length_mm)) +
  geom_point(aes(color = species),
             size = 3, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, aes(color = species)) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper and bill length",
       caption = "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)", y = "Bill length (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = "bottom",
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face = "italic"),
        plot.caption.position = "plot")
```
