---
title: "Data Wrangling"
subtitle: "Iteration"
css: custom_report.css
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme: default
    highlight: tango
    fig_captions: no
    number_sections: false
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
  knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
  options(scipen=999) # turns off scientific notation

# define colors
  glhlth562_colors <- c(
    `blue`       = "#1f9ac9",
    `yellow`     = "#ffd966",
    `black`      = "#000000",
    `lightgrey`  = "#D3D3D3",
    `darkgrey`   = "#444444",
    `grey`       = "#808080"
  )

# load packages
  library(tidyverse)
  library(palmerpenguins)
  library(broom)
```

# Iteration



## It might even be in your present

Take this example, but imagine you want 3 separate plots, not facets.

```{r}
ggplot(penguins, 
       aes(x = bill_length_mm)) +
  geom_histogram() +
  facet_wrap(~species, nrow = 3) + 
  theme_bw() +
  theme(plot.title = 
          element_text(face="bold"))
```

## Copy/paste with few changes

```{r}
p1 <- penguins %>% filter(species == "Adelie") %>%
  ggplot(aes(x = bill_length_mm)) + geom_histogram() + 
  labs(x = "Bill length (mm)", y = NULL,
       title = "Distribution of bill length among Adelie penguins") +
  theme_bw() + theme(plot.title = element_text(face="bold"))

p2 <- penguins %>% filter(species == "Gentoo") %>%
  ggplot(aes(x = bill_length_mm)) + geom_histogram() +
  labs(x = "Bill length (mm)", y = NULL,
       title = "Distribution of bill length among Gentoo penguins") +
  theme_bw() + theme(plot.title = element_text(face="bold"))

p3 <- penguins %>% filter(species == "Chinstrap") %>%
  ggplot(aes(x = bill_length_mm)) + geom_histogram() + 
  labs(x = "Bill length (mm)", y = NULL,
       title = "Distribution of bill length among Chinstrap penguins") +
  theme_bw() + theme(plot.title = element_text(face="bold"))
```

## Here's a function to create these plots

```{r}
plot_bill_length <- function(penguin_species) {
  
  penguins %>%
    filter(species == penguin_species) %>%
    ggplot(aes(x = bill_length_mm)) +
    geom_histogram() + 
    labs(x = "Bill length (mm)",
         y = NULL,
         title = glue::glue("Distribution of bill length among",
                            penguin_species,
                            "penguins",
                            .sep = " ")) +
    theme_bw() +
      theme(plot.title = element_text(face="bold"))
}
```

## Give into writing functions

```{r}
p1 <- plot_bill_length(penguin_species = "Adelie")
p2 <- plot_bill_length(penguin_species = "Gentoo")
p3 <- plot_bill_length(penguin_species = "Chinstrap")
```


## Let's break this down

```{r, eval=FALSE}
plot_bill_length <- function(penguin_species) {
  
  penguins %>%
    filter(species == penguin_species) %>%
    ggplot(aes(x = bill_length_mm)) +
    geom_histogram() + 
    labs(x = "Bill length (mm)",
         y = NULL,
         title = glue::glue("Distribution of bill length among",
                            penguin_species,
                            "penguins",
                            .sep = " ")) +
    theme_bw() +
      theme(plot.title = element_text(face="bold"))
}
```

## For loops + functions

```{r}
for (`s` in c("Adelie", "Gentoo", "Chinstrap")) {
  print(plot_bill_length(penguin_species = `s`))
  # alternatively ggsave(filename = glue::glue(s, ".png"))
}
```

## Here's a basic for loop setup

Compute the mean of every column in mtcars

```{r}
output <- vector("double", ncol(mtcars))
names(output) <- names(mtcars)
for (i in names(mtcars)) {
  output[i] <- mean(mtcars[[i]])
}

output
```

## Evergreen sidenote: There are usually many pathways from A to B

```{r}
mtcars %>%
  summarise_all(mean)
```

## Here's an example


```{r}
`map`(c("Adelie", "Gentoo", "Chinstrap"), ~ `plot_bill_length(.)`)
```

## Anonymous functions

```{r}
map_dbl(mtcars, ~ length(unique(.x)))
```

### Passing arguments with `...`

```{r}
x <- list(1:5, c(1:10, NA))
x
map_dbl(x, mean, na.rm = TRUE)
```


## A typical use case

```{r}
mtcars %>%
  group_by(cyl) %>%
  nest() %>%
  mutate(results = map(.x = data, ~ lm(mpg ~ wt, data = .x) %>% tidy())) %>%
  unnest(cols = results) %>%
  filter(term=="wt")
```

