---
title: "The Office 1"
author: "Your name"
output: github_document
---

```{r load-packages, message = FALSE}
# credit to https://datasciencebox.org/index.html
library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
```

In this two-part exercise we will use `theoffice` data from the [**schrute**](https://bradlindblad.github.io/schrute/) package to predict IMDB scores for episodes of The Office. In the first part, we'll work together to prepare the data, specify a model, and create a recipe. In the second part, you will fit and evaluate the model on your own for homework.

```{r}
glimpse(theoffice)
```

Fix `air_date` for later use.

```{r}
theoffice <- theoffice %>%
  mutate(air_date = ymd(as.character(air_date)))
```

We will

-   engineer features based on episode scripts
-   train a model
-   perform cross validation
-   make predictions

```{r}
theoffice %>%
  distinct(season, episode)
```

### Exercise 1 - Calculate the percentage of lines spoken by Jim, Pam, Michael, and Dwight for each episode of The Office.

```{r lines}

```

### Exercise 2 - Identify episodes that touch on Halloween, Valentine's Day, and Christmas.

```{r special-episodes}

```

### Exercise 3 - Put together a modeling dataset that includes features you've engineered. Also add an indicator variable called `michael` which takes the value `1` if Michael Scott (Steve Carrell) was there, and `0` if not. Note: Michael Scott (Steve Carrell) left the show at the end of Season 7.

```{r office-df}
  
```

### Exercise 4 - Split the data into training (75%) and testing (25%).

```{r split}

```

### Exercise 5 - Specify a linear regression model.

```{r model}

```

### Exercise 6 - Create a recipe that updates the role of `episode_name` to not be a predictor, removes `air_date` as a predictor, uses `season` as a factor, and removes all zero variance predictors.

```{r recipe}

```

### Exercise 7 - Build a workflow for fitting the model specified earlier and using the recipe you developed to preprocess the data.

```{r workflow}

```
